---
title: "R Notebook"
output: html_notebook
---

#загружаем библиотеки
```{r}
library(knitr)
library(xtable) # table
library(mixOmics)
library(sva) # ComBat
library(ggplot2) # PCA sample plot with density
library(gridExtra) # PCA sample plot with density
library(limma) # removeBatchEffect (LIMMA)
library(vegan) # RDA
library(AgiMicroRna) # RLE plot
library(cluster) # silhouette coefficient
library(variancePartition) # variance calculation
library(pvca) # PVCA
library(pheatmap) # heatmap
library(ruv) # RUVIII
library(lmerTest) # lmer
library(bapred)
library(FactoMineR)# FAbatch
library(phyloseq)
library(devtools)
library(tidyr)
library(ggpubr)
library(phyloseq)
library(plyr)
library(dplyr)
library(vegan)
library(pheatmap)
library(DESeq2)
library(ggpubr)
library(matrixStats)
library(tibble)
library(HTSSIP)
source('/Users/basilews/phylo/helpers.R')
#их фаил с доп функциями
source('/Users/basilews/phylo/batc_panc/DATA/exp/helpers.R')
```

#загружаем данные
```{r}
seqtab <- readRDS('/Users/basilews/phylo/batc_panc/DATA/panc1/seqtab_nochim__20.rds')
seqtab_1 <- readRDS('/Users/basilews/phylo/batc_panc/DATA/panc2/seqtab_nochim__20.rds')
taxa_1 <- readRDS('/Users/basilews/phylo/batc_panc/DATA/panc2/taxa_20.rds')
meta_1 <- read.csv('/Users/basilews/phylo/batc_panc/DATA/panc2/panc_meta2b.csv',row.names = 1)
rownames(seqtab) <- gsub('.{3}$', '', sapply(strsplit(rownames(seqtab), ".", fixed = TRUE), `[`, 1))
rownames(seqtab_1) <- gsub('.{3}$', '', sapply(strsplit(rownames(seqtab_1), ".", fixed = TRUE), `[`, 1))
taxa   <- readRDS('/Users/basilews/phylo/batc_panc/DATA/panc1/taxa_20.rds')
meta <- read.csv('/Users/basilews/phylo/batc_panc/DATA/panc1/panc_metab.csv',row.names = 1,sep=';')
colnames(meta)[which(names(meta ) == "Batch")] <- "batch"
colnames(meta_1)[which(names(meta_1 ) == "batch")] <- "batch"
```

```{r}
ps <- phyloseq(otu_table(seqtab, taxa_are_rows=FALSE)
               ,tax_table(taxa) 
               ,sample_data(meta)
)
n <- as.data.frame(sample_sums(ps))
c <- c('P005EP','P005KR','P010','P010EP','P008','P008EP')
'%ni%' <- Negate('%in%')
my_subset <- subset(ps@sam_data, rownames(ps@sam_data) %ni% c)
ps <- phyloseq(otu_table(seqtab, taxa_are_rows=FALSE)
               ,tax_table(taxa) 
               ,sample_data(my_subset)
)
ps_2 <- phyloseq(otu_table(seqtab_1, taxa_are_rows=FALSE)
                 ,tax_table(taxa_1) 
                 ,sample_data(meta_1)
)


taxa_name_func <- function(x) {paste0(x['Phylum'], '.c__' , substr(x['Class'], 1, 3), '.o__' , substr(x['Order'], 1, 5), '.f__' , substr(x['Family'], 1, 3), '.g__' , gsub('/', '', gsub("-", "", x['Genus'])))}

# MERGE THEM
#CREATE NEW PHYLOSEQ OBJECT FOR PANCREATITIS1
ps_obj <- ps
otu_matrix <- as(otu_table(ps_obj), 'matrix')
taxa_matrix <- as(tax_table(ps_obj), 'matrix')
taxa_matrix_good_names <- apply(taxa_matrix, MARGIN=1, taxa_name_func)
# ps_meta <- as(sample_data(ps_obj), 'matrix')
colnames(otu_matrix) <- taxa_matrix_good_names
rownames(taxa_matrix) <- taxa_matrix_good_names

count_mtrx_genus <- otu_matrix[, !duplicated(colnames(otu_matrix))]
taxa_genus <- taxa_matrix[!duplicated(rownames(taxa_matrix)),]

ps <- phyloseq(otu_table(count_mtrx_genus, taxa_are_rows=FALSE)
               ,tax_table(taxa_genus) 
               ,sample_data(ps@sam_data)
)


# CREATE NEW PHYLOSEQ OBJECT FOR panc2
ps_obj <- ps_2
otu_matrix <- as(otu_table(ps_obj), 'matrix')
taxa_matrix <- as(tax_table(ps_obj), 'matrix')
taxa_matrix_good_names <- apply(taxa_matrix, MARGIN=1, taxa_name_func)
# ps_meta <- as(sample_data(ps_obj), 'matrix')
colnames(otu_matrix) <- taxa_matrix_good_names
rownames(taxa_matrix) <- taxa_matrix_good_names

count_mtrx_genus <- otu_matrix[, !duplicated(colnames(otu_matrix))]
taxa_genus <- taxa_matrix[!duplicated(rownames(taxa_matrix)),]

ps_2 <- phyloseq(otu_table(count_mtrx_genus, taxa_are_rows=FALSE)
                 ,tax_table(taxa_genus) 
                 ,sample_data(ps_2@sam_data)
)
ps_all <- merge_phyloseq(ps_2,ps)


```
#предобработка данных
```{r}
otu.panc <- as(otu_table(ps_all), 'matrix')
dim(otu.panc)
otu.panc.keep <- which(colSums(otu.panc)*100/(sum(colSums(otu.panc))) > 0.01)
otu.panc.keep <- otu.panc[, otu.panc.keep]
dim(otu.panc.keep)
#otu.panc <- otu.panc + 0.01
otu.panc.keep <- otu.panc.keep + 1
panc.clr <- logratio.transfo(otu.panc.keep, logratio = 'CLR')
class(panc.clr) <- 'matrix' 
```

#Pheatmap
```{r}
ps_all_clr <- ps_all
otu_table(ps_all_clr) <- otu_table(panc.clr,taxa_are_rows = F)
ps_obj <- ps_all_clr
#ps_obj <- filter_taxa(ps_genus, function(x){sum(x > 0) > 10}, prune = TRUE)
# prepare metadata
ps_obj_meta <- sample_data(ps_obj)
cols_of_interest =c('group','batch')
meta_for_heatmap <- ps_obj_meta[, (names(ps_obj_meta) %in% cols_of_interest)]
meta_for_heatmap <- as(meta_for_heatmap, 'data.frame')
ordered_meta <- meta_for_heatmap[order(meta_for_heatmap$group,meta_for_heatmap$batch),]

 pheatmap( 
  t((ps_all_clr@otu_table)), 
  cluster_rows = F, 
  cluster_cols = T,
  fontsize_row = 4, fontsize_col = 5,
  fontsize = 12,
  clustering_distance_rows = 'euclidean',
  clustering_method = 'ward.D',
  treeheight_row = 15,
  annotation_col = meta_for_heatmap,
  annotation_colors = annotation_colors,
  # col= hmcols, 
  # breaks = bk,
  main='before correction')
```

#RLE before batch correction by groups
```{r}

panc.batch_l <- panc.batch[panc.gr == 'легкая степень']
panc.batch_t <- panc.batch[panc.gr == 'тяжелая степень'] 
panc.batch_h <- panc.batch[panc.gr == 'хр панкреатит'] 
panc.before_l <- panc.clr[panc.gr == 'легкая степень', ]
panc.before_t <- panc.clr[panc.gr == 'тяжелая степень', ] 
panc.before_h <- panc.clr[panc.gr == 'хр панкреатит', ] 

par(mfrow = c(1,3))
RleMicroRna2(object = t(panc.before_l), batch = panc.batch_l, 
             maintitle = 'Легкая степень')

RleMicroRna2(object = t(panc.before_t), batch = panc.batch_t, 
             maintitle = 'Тяжелая с тепень')
RleMicroRna2(object = t(panc.before_h), batch = panc.batch_h, 
             maintitle = 'Хронический панкреатит')
```

#Correcting batch effects
```{r}

panc.mod <- model.matrix( ~ panc.gr)
#BMC (batch mean centering)
# Sponge data
panc.b1 <- scale(panc.clr[panc.batch == 1, ], center = TRUE, scale = FALSE)
panc.b2 <- scale(panc.clr[panc.batch == 2, ], center = TRUE, scale = FALSE)
panc.bmc <- rbind(panc.b1, panc.b2)
panc.bmc <- panc.bmc[rownames(panc.clr), ]

#ComBat
panc.combat <- t(ComBat(t(panc.clr), batch = panc.batch, 
                          mod = panc.mod, par.prior = F, prior.plots = F))

#removeBatchEffect

panc.limma <- t(removeBatchEffect(t(panc.clr), batch = panc.batch, 
                                    design = panc.mod))
#FAbatch 
panc.gr.f <- recode_factor(panc.gr, "легкая степень" = "тяжелая степень")

panc.fabatch.obj <- fabatch(x = panc.clr, 
                              y = as.factor(as.numeric(panc.gr.f)), 
                              batch = panc.batch)
panc.fabatch <- panc.fabatch.obj$xadj

#Percentile normalisation
panc.tss <- t(apply(otu.panc.keep, 1, function(x){x/sum(x)}))
panc.percentile <- percentile_norm(data = panc.tss, batch = panc.batch,trt = panc.gr)
#SVD-based method

panc.sd <- apply(panc.clr, 2, sd) # calculate standard deviation
panc.mean <- apply(panc.clr, 2, mean) # calculate mean
panc.X <- scale(panc.clr, center = T, scale = T) # center and scale

panc.m <- crossprod(panc.X) # generate a square matrix
panc.m.svd <- svd(panc.m) # SVD 
# extract 1st singular vectors
panc.a1 <- panc.m.svd$u[ ,1] 
panc.b1 <- panc.m.svd$v[ ,1]

# deflate component 1 from the data
panc.t1 <- panc.X %*% panc.a1 / drop(sqrt(crossprod(panc.a1)))
panc.c1 <- crossprod(panc.X, panc.t1) / drop(crossprod(panc.t1))
panc.svd.defl.matrix1  <- panc.X - panc.t1 %*% t(panc.c1)

# add back mean and standard deviation
panc.svd <- panc.svd.defl.matrix1
panc.svd[1:nrow(panc.svd), 1:ncol(panc.svd)] = NA
for(i in 1:ncol(panc.svd.defl.matrix1)){
  for(j in 1:nrow(panc.svd.defl.matrix1)){
    panc.svd[j,i] = panc.svd.defl.matrix1[j,i]*panc.sd[i] + panc.mean[i]
  }
}

```

#Methods evaluation
#Principal component analysis (PCA) with density plot per component
```{r}

panc.pca.before <- pca(panc.clr, ncomp = 3)
panc.pca.bmc <- pca(panc.bmc, ncomp = 3)
panc.pca.combat <- pca(panc.combat, ncomp = 3)
panc.pca.limma <- pca(panc.limma, ncomp = 3)
panc.pca.percentile <- pca(panc.percentile, ncomp = 3)
panc.pca.svd <- pca(panc.svd, ncomp = 3)
panc.pca.fa <- pca(panc.fabatch, ncomp = 3)

panc.pca.plot.fa = Scatter_Density(data = panc.pca.fa$variates$X, batch = panc.batch, 
                                       trt = panc.gr, expl.var = panc.pca.fa$explained_variance,
                                       xlim = c(-20,20), ylim = c(-20,20),
                                       batch.legend.title = 'Seq data (batch)', 
                                       trt.legend.title = 'Group (trt)', 
                                       title = 'Fabatch batch effect correction')
panc.pca.plot.before = Scatter_Density(data = panc.pca.before$variates$X, batch = panc.batch, 
                                         trt = panc.gr, expl.var = panc.pca.before$explained_variance,
                                       xlim = c(-20,20), ylim = c(-20,20),
                                         batch.legend.title = 'Seq data (batch)', 
                                         trt.legend.title = 'Group (trt)', 
                                         title = 'Before batch effect correction')
panc.pca.plot.bmc = Scatter_Density(data = panc.pca.bmc$variates$X, batch = panc.batch, 
                                      trt = panc.gr, expl.var = panc.pca.bmc$explained_variance, 
                                      xlim = c(-20,20), ylim = c(-20,20), 
                                      batch.legend.title = 'Seq data (batch)', 
                                      trt.legend.title = 'Group (trt)', 
                                      title = 'BMC Correction')
panc.pca.plot.combat = Scatter_Density(data = panc.pca.combat$variates$X, batch = panc.batch, 
                                         trt = panc.gr, expl.var = panc.pca.combat$explained_variance, 
                                         xlim = c(-20,20), ylim = c(-20,20), 
                                       batch.legend.title = 'Seq data (batch)', 
                                       trt.legend.title = 'Group (trt)', 
                                       title = 'Combat Correction')
panc.pca.plot.limma = Scatter_Density(data = panc.pca.limma$variates$X, batch = panc.batch, 
                                        trt = panc.gr, expl.var = panc.pca.limma$explained_variance, 
                                     xlim = c(-20,20), ylim = c(-20,20), 
                                     batch.legend.title = 'Seq data (batch)', 
                                     trt.legend.title = 'Group (trt)', 
                                     title = 'Limma(rBE) Correction')
grid.arrange(panc.pca.plot.before, panc.pca.plot.bmc, 
             panc.pca.plot.combat, panc.pca.plot.limma, ncol = 2)

```
```{r}
panc.pca.plot.percentile = Scatter_Density(data = panc.pca.percentile$variates$X, batch = panc.batch, 
                                             trt = panc.gr, expl.var = panc.pca.percentile$explained_variance, 
                                            xlim = c(-20,20), ylim = c(-20,20), 
                                           batch.legend.title = 'Seq data (batch)', 
                                           trt.legend.title = 'Group (trt)', 
                                           title = 'Percentile Correction')
panc.pca.plot.svd = Scatter_Density(data = panc.pca.svd$variates$X, batch = panc.batch, 
                                                      trt = panc.gr, expl.var = panc.pca.svd$explained_variance, 
                                                      xlim = c(-20,20), ylim = c(-20,20), 
                                                      batch.legend.title = 'Seq data (batch)', 
                                                      trt.legend.title = 'Group (trt)', 
                                                      title = 'SVD Correction')
grid.arrange(panc.pca.plot.before, panc.pca.plot.percentile, 
             panc.pca.plot.svd,panc.pca.plot.fa, ncol = 2)

```


#box plots for OTU
```{r}
panc.before.df <- data.frame(value = panc.clr[ ,9], batch = panc.batch)
panc.boxplot.before <- box_plot_fun(data = panc.before.df, 
                                      x = panc.before.df$batch,
                                      y = panc.before.df$value, 
                                      title = 'OTU9 - before', 
                                      batch.legend.title = 'Seq data (batch)')

panc.bmc.df <- data.frame(value = panc.bmc[ ,9], batch = panc.batch)
panc.boxplot.bmc <- box_plot_fun(data = panc.bmc.df, 
                                   x = panc.bmc.df$batch,
                                   y = panc.bmc.df$value, 
                                   title = 'OTU9 - BMC', 
                                   batch.legend.title = 'Seq data (batch)')


panc.combat.df <- data.frame(value = panc.combat[ ,9], batch = panc.batch)
panc.boxplot.combat <- box_plot_fun(data = panc.combat.df, 
                                      x = panc.combat.df$batch, 
                                      y = panc.combat.df$value, 
                                      title = 'OTU9 - ComBat',
                                      batch.legend.title = 'Seq data (batch)')


panc.limma.df <- data.frame(value = panc.limma[ ,9], batch = panc.batch)
panc.boxplot.limma <- box_plot_fun(data = panc.limma.df, 
                                     x = panc.limma.df$batch, 
                                     y = panc.limma.df$value, 
                                     title = 'OTU9 - rBE', 
                                     batch.legend.title = 'Data seq (batch)')

panc.percentile.df <- data.frame(value = panc.percentile[ ,9], batch = panc.batch)
panc.boxplot.percentile <- box_plot_fun(data = panc.percentile.df, 
                                          x = panc.percentile.df$batch, 
                                          y = panc.percentile.df$value, 
                                          title = 'OTU9 - PN ', 
                                          batch.legend.title = 'Seq data(batch)')

panc.svd.df <- data.frame(value = panc.svd[ ,9], batch = panc.batch)
panc.boxplot.svd <- box_plot_fun(data = panc.svd.df, 
                                   x = panc.svd.df$batch, 
                                   y = panc.svd.df$value, 
                                   title = 'OTU9 - SVD ', 
                                   batch.legend.title = 'Seq data (batch)')

panc.fabatch.df <- data.frame(value = panc.fabatch[ ,9], batch = panc.batch)
panc.boxplot.fabatch <- box_plot_fun(data = panc.svd.df, 
                                   x = panc.fabatch.df$batch, 
                                   y = panc.fabatch.df$value, 
                                   title = 'OTU9 - Fabatch ', 
                                   batch.legend.title = 'Seq data (batch)')


grid.arrange(panc.boxplot.before, panc.boxplot.bmc, 
             panc.boxplot.combat, panc.boxplot.limma, ncol = 2)
grid.arrange(panc.boxplot.before, panc.boxplot.percentile, 
             panc.boxplot.svd,panc.boxplot.fabatch, ncol = 2)
```

#density plots
```{r}

# before
panc.dens.before <- ggplot(panc.before.df, aes(x = value, fill = batch)) + 
  geom_density(alpha = 0.5) + scale_fill_manual(values = color.mixo(1:10)) + 
  labs(title = 'OTU9 - before ', x = 'Value', fill = 'Seq data (batch)') + 
  theme_bw() + theme(plot.title = element_text(hjust = 0.5), 
                     panel.grid = element_blank())

# BMC
panc.dens.bmc <- ggplot(panc.bmc.df, aes(x = value, fill = batch)) + 
  geom_density(alpha = 0.5) + scale_fill_manual(values = color.mixo(1:10)) + 
  labs(title = 'OTU9 - BMC', x = 'Value', fill = 'Seq data  (batch)') + 
  theme_bw() + theme(plot.title = element_text(hjust = 0.5), 
                     panel.grid = element_blank())


# ComBat
panc.dens.combat <- ggplot(panc.combat.df, aes(x = value, fill = batch)) + 
  geom_density(alpha = 0.5) + scale_fill_manual(values = color.mixo(1:10)) + 
  labs(title = 'OTU9 - ComBat ', x = 'Value', fill = 'Seq data  (batch)') + 
  theme_bw() + theme(plot.title = element_text(hjust = 0.5), 
                     panel.grid = element_blank())


# removeBatchEffect 
panc.dens.limma <- ggplot(panc.limma.df, aes(x = value, fill = batch)) + 
  geom_density(alpha = 0.5) + scale_fill_manual(values = color.mixo(1:10)) + 
  labs(title = 'OTU9 - rBE ', x = 'Value', fill = 'Seq data  (batch)') + 
  theme_bw() + theme(plot.title = element_text(hjust = 0.5), 
                     panel.grid = element_blank())


# percentile normal
panc.dens.percentile <- ggplot(panc.percentile.df, aes(x = value, fill = batch)) + 
  geom_density(alpha = 0.5) + scale_fill_manual(values = color.mixo(1:10)) + 
  labs(title = 'OTU9 - PN ', x = 'Value', fill = 'Seq data  (batch)') + 
  theme_bw() + theme(plot.title = element_text(hjust = 0.5), 
                     panel.grid = element_blank())


# SVD
panc.dens.svd <- ggplot(panc.svd.df, aes(x = value, fill = batch)) + 
  geom_density(alpha = 0.5) + scale_fill_manual(values = color.mixo(1:10)) + 
  labs(title = 'OTU9 - SVD', x = 'Value', fill = 'Seq data  (batch)') + 
  theme_bw() + theme(plot.title = element_text(hjust = 0.5), 
                     panel.grid = element_blank())
#fabatch
panc.dens.fabatch <- ggplot(panc.fabatch.df, aes(x = value, fill = batch)) + 
  geom_density(alpha = 0.5) + scale_fill_manual(values = color.mixo(1:10)) + 
  labs(title = 'OTU9 - Fabatch', x = 'Value', fill = 'Seq data  (batch)') + 
  theme_bw() + theme(plot.title = element_text(hjust = 0.5), 
                     panel.grid = element_blank())

grid.arrange(panc.dens.before, panc.dens.bmc, 
             panc.dens.combat, panc.dens.limma, panc.dens.percentile, 
             panc.dens.svd, panc.dens.fabatch,ncol = 2)


```
# P-values
```{r}

panc.lm.before <- lm(panc.clr[ ,9] ~ panc.gr + panc.batch)
summary(panc.lm.before)

panc.lm.bmc <- lm(panc.bmc[ ,9] ~ panc.gr + panc.batch)
summary(panc.lm.bmc)  
panc.lm.combat <- lm(panc.combat[ ,9] ~ panc.gr + panc.batch)
summary(panc.lm.combat)

panc.lm.limma <- lm(panc.limma[ ,9] ~ panc.gr + panc.batch)
summary(panc.lm.limma)

panc.lm.percentile <- lm(panc.percentile[ ,9] ~ panc.gr + panc.batch)
summary(panc.lm.percentile)

panc.lm.svd <- lm(panc.svd[ ,9] ~ panc.gr + panc.batch)
summary(panc.lm.svd)

panc.lm.fabatch <- lm(panc.fabatch[ ,9] ~ panc.gr + panc.batch)
summary(panc.lm.svd)
```

#RLE
```{r}
# BMC
panc.batch_l <- panc.batch[panc.gr == 'легкая степень']
panc.batch_t <- panc.batch[panc.gr == 'тяжелая степень'] 
panc.batch_h <- panc.batch[panc.gr == 'хр панкреатит'] 
panc.before_l <- panc.clr[panc.gr == 'легкая степень', ]
panc.before_t <- panc.clr[panc.gr == 'тяжелая степень', ] 
panc.before_h <- panc.clr[panc.gr == 'хр панкреатит', ] 


panc.bmc_l <- panc.bmc[panc.gr == 'легкая степень', ]
panc.bmc_t <- panc.bmc[panc.gr == 'тяжелая степень', ] 
panc.bmc_h <- panc.bmc[panc.gr == 'хр панкреатит',]

# ComBat
panc.combat_l <- panc.combat[panc.gr == 'легкая степень', ]
panc.combat_t <- panc.combat[panc.gr == 'тяжелая степень', ] 
panc.combat_h <- panc.combat[panc.gr == 'хр панкреатит',]

# rBE
panc.limma_l <- panc.limma[panc.gr == 'легкая степень', ]
panc.limma_t <- panc.limma[panc.gr == 'тяжелая степень', ] 
panc.limma_h <- panc.limma[panc.gr == 'хр панкреатит',] 

# PN
panc.percentile_l <- panc.percentile[panc.gr == 'легкая степень', ]
panc.percentile_t <- panc.percentile[panc.gr == 'тяжелая степень', ] 
panc.percentile_h <- panc.percentile[panc.gr == 'хр панкреатит',] 

# SVD
panc.svd_l <- panc.svd[panc.gr == 'легкая степень', ]
panc.svd_t <- panc.svd[panc.gr == 'тяжелая степень', ] 
panc.svd_h <- panc.svd[panc.gr == 'хр панкреатит',] 

#FAbatch
panc.fa_t <- panc.fabatch[panc.gr == 'тяжелая степень', ] 
panc.fa_l <- panc.fabatch[panc.gr == 'легкая степень', ] 
panc.fa_h <- panc.fabatch[panc.gr == 'хр панкреатит',] 

```
#Rle
```{r}
par(mfrow = c(3,3), mai = c(0.4,0.6,0.3,0.1))

RleMicroRna2(object = t(panc.before_l), batch = panc.batch_l, 
             maintitle = 'Panc: before (легкая степень)', title.cex = 1)

RleMicroRna2(object = t(panc.bmc_l), batch = panc.batch_l, 
             maintitle = 'Panc: bmc (легкая степень)', title.cex = 1)

RleMicroRna2(object = t(panc.combat_l), batch = panc.batch_l,
             maintitle = 'Panc: combat (легкая степень)', title.cex = 1)

RleMicroRna2(object = t(panc.limma_l ), batch = panc.batch_l, 
             maintitle = 'Panc: limma (легкая степень)', title.cex = 1)

RleMicroRna2(object = t(panc.percentile_l), batch = panc.batch_l, 
             maintitle = 'Panc: PN (легкая степень)', title.cex = 1)

RleMicroRna2(object = t(panc.svd_l), batch = panc.batch_l, 
             maintitle = 'Panc: svd (легкая степень)', title.cex = 1)
RleMicroRna2(object = t(panc.fa_l), batch = panc.batch_t, 
             maintitle = 'Panc: Fabacth (легкая степень)', title.cex = 1)
```
#Rle
```{r}
par(mfrow = c(3,3), mai = c(0.4,0.6,0.3,0.1))

RleMicroRna2(object = t(panc.before_t), batch = panc.batch_t, 
             maintitle = 'Panc: before (тяжелая степень)', title.cex = 1)

RleMicroRna2(object = t(panc.bmc_t), batch = panc.batch_t, 
             maintitle = 'Panc: bmc (тяжелая степень)', title.cex = 1)

RleMicroRna2(object = t(panc.combat_t), batch = panc.batch_t, 
             maintitle = 'Panc: combat (тяжелая степень)', title.cex = 1)

RleMicroRna2(object = t(panc.limma_t ), batch = panc.batch_t, 
             maintitle = 'Panc: limma (тяжелая степень)', title.cex = 1)

RleMicroRna2(object = t(panc.percentile_t),batch = panc.batch_t, 
             maintitle = 'Panc: percentile (тяжелая степень)', title.cex = 1)

RleMicroRna2(object = t(panc.svd_t), batch = panc.batch_t, 
             maintitle = 'Panc: svd (тяжелая степень)', title.cex = 1)
RleMicroRna2(object = t(panc.fa_t), batch = panc.batch_t, 
             maintitle = 'Panc: Fabacth (тяжелая степень)', title.cex = 1)
```
#Rle
```{r}
par(mfrow = c(3,3), mai = c(0.4,0.6,0.3,0.1))
RleMicroRna2(object = t(panc.before_h), batch = panc.batch_h, 
             maintitle = 'Panc: before (хр панткреатит)', title.cex = 1)

RleMicroRna2(object = t(panc.bmc_h), batch = panc.batch_h, 
             maintitle = 'Panc: bmc (хр панткреатит)', title.cex = 1)

RleMicroRna2(object = t(panc.combat_h), batch = panc.batch_h, 
             maintitle = 'Panc: combat (хр панткреатит)', title.cex = 1)

RleMicroRna2(object = t(panc.limma_h ), batch = panc.batch_h, 
             maintitle = 'Panc: limma (хр панткреатит)', title.cex = 1)

RleMicroRna2(object = t(panc.percentile_h),batch = panc.batch_h, 
             maintitle = 'Panc: percentile (хр панткреатит)', title.cex = 1)

RleMicroRna2(object = t(panc.svd_h), batch = panc.batch_h, 
             maintitle = 'Panc: svd (хр панткреатит)', title.cex = 1)

RleMicroRna2(object = t(panc.fa_h), batch = panc.batch_t, 
             maintitle = 'Panc: Fabacth (хр панкреатит)', title.cex = 1)

```
#Linear model per variable
```{r}
#Linear model per variable

panc.form <- ~ panc.gr + panc.batch
panc.info <- as.data.frame(cbind(rownames(panc.clr),panc.gr, panc.batch))
rownames(panc.info) <- rownames(panc.clr)

# before
panc.varPart.before <- fitExtractVarPartModel(exprObj = t(panc.clr), 
                                                formula = panc.form, 
                                                data = panc.info)

# BMC
panc.varPart.bmc <- fitExtractVarPartModel(exprObj = t(panc.bmc), 
                                             formula = panc.form, 
                                             data = panc.info)
#fabacth
panc.varPart.fab <- fitExtractVarPartModel(exprObj = t(panc.fabatch), 
                                           formula = panc.form, 
                                           data = panc.info)
# combat
panc.varPart.combat <- fitExtractVarPartModel(exprObj = t(panc.combat), 
                                                formula = panc.form, 
                                                data = panc.info)

# removeBatchEffect
panc.varPart.limma <- fitExtractVarPartModel(exprObj = t(panc.limma), 
                                               formula = panc.form, 
                                               data = panc.info)

# percentile normalisation
panc.varPart.percentile <- fitExtractVarPartModel(exprObj = t(panc.percentile), 
                                                    formula = panc.form, 
                                                    data = panc.info)

# svd
panc.varPart.svd <- fitExtractVarPartModel(exprObj = t(panc.svd), 
                                             formula = panc.form, 
                                             data = panc.info)
# extract the variance of trt and batch
# before
panc.varmat.before <- as.matrix(panc.varPart.before[ ,1:2])
# BMC
panc.varmat.bmc <- as.matrix(panc.varPart.bmc[ ,1:2])
# ComBat
panc.varmat.combat <- as.matrix(panc.varPart.combat[ ,1:2])
# removeBatchEffect
panc.varmat.limma <- as.matrix(panc.varPart.limma[ ,1:2])
# percentile normalisation
panc.varmat.percentile <- as.matrix(panc.varPart.percentile[ ,1:2])
# SVD
panc.varmat.svd <- as.matrix(panc.varPart.svd[ ,1:2])
#fabatch
panc.varmat.fab <- as.matrix(panc.varPart.fab[ ,1:2])
# merge results
panc.variance <- c(as.vector(panc.varmat.before), as.vector(panc.varmat.bmc),
                     as.vector(panc.varmat.combat), as.vector(panc.varmat.limma),
                     as.vector(panc.varmat.percentile), as.vector(panc.varmat.svd),as.vector(panc.varmat.fab))

# add batch, trt and methods info
panc.variance <- cbind(variance = panc.variance, 
                         Type = rep(c('Group', 'Batch'), each = ncol(panc.clr)),
                         method = rep(c('Before', 'BMC', 'ComBat', 'rBE', 
                                        'PN', 'SVD','Fabatch'), each = 2*ncol(panc.clr)))
# reorder levels  
panc.variance <- as.data.frame(panc.variance)
panc.variance$method <- factor(panc.variance$method, 
                                 levels = unique(panc.variance$method))
panc.variance$variance <- as.numeric(as.character(panc.variance$variance))

ggplot(panc.variance, aes(x = Type, y = variance, fill = Type)) + 
  geom_boxplot() + facet_grid(cols = vars(method)) + theme_bw() + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1), 
        strip.text = element_text(size = 12), panel.grid = element_blank(), 
        axis.text = element_text(size = 12), axis.title = element_text(size = 15), 
        legend.title = element_text(size = 15), legend.text = element_text(size = 12)) + 
  labs(x = 'Type', y = 'Proportion Variance', name = 'Type') + ylim(0,1)


panc.b.coeff <- c()
for(i in 1:ncol(panc.clr)){
  res <- lm(panc.clr[ ,i] ~ panc.gr + panc.batch)
  sum.res <- summary(res)
  panc.b.coeff[i] <- sum.res$coefficients[3,1]
}

```
#Pheatmap
#Combat
```{r}
ps_combat<- ps_all
otu_table(ps_combat) <- otu_table(panc.combat,taxa_are_rows = F)
ps_obj <- ps_combat
# prepare metadata
ps_obj_meta <- sample_data(ps_obj)
cols_of_interest =c('group','batch')
meta_for_heatmap <- ps_obj_meta[, (names(ps_obj_meta) %in% cols_of_interest)]
meta_for_heatmap <- as(meta_for_heatmap, 'data.frame')
ordered_meta <- meta_for_heatmap[order(meta_for_heatmap$group,meta_for_heatmap$batch),]
pheatmap( 
  t((ps_combat@otu_table)), 
  cluster_rows = F, 
  cluster_cols = T,
  fontsize_row = 4, fontsize_col = 5,
  fontsize = 12,
  #clustering_distance_rows = 'euclidean',
  #clustering_method = 'ward.D',
  treeheight_row = 15,
  annotation_col = meta_for_heatmap,
  annotation_colors = annotation_colors,
  #height = 40,
  # col= hmcols, 
  # breaks = bk,
  main='Combat correction')

```
#BMC
```{r}
ps_bmc<- ps_all
otu_table(ps_bmc) <- otu_table(panc.bmc,taxa_are_rows = F)
ps_obj <- ps_bmc
ps_obj_meta <- sample_data(ps_obj)
cols_of_interest =c('group','batch')
meta_for_heatmap <- ps_obj_meta[, (names(ps_obj_meta) %in% cols_of_interest)]
meta_for_heatmap <- as(meta_for_heatmap, 'data.frame')
ordered_meta <- meta_for_heatmap[order(meta_for_heatmap$group,meta_for_heatmap$batch),]
pheatmap( 
  t((ps_bmc@otu_table)), 
  cluster_rows = F, 
  cluster_cols = T,
  fontsize_row = 4, fontsize_col = 5,
  fontsize = 12,
  clustering_distance_rows = 'euclidean',
  clustering_method = 'ward.D',
  treeheight_row = 15,
  annotation_col = meta_for_heatmap,
  annotation_colors = annotation_colors,
  height = 40,
  # col= hmcols, 
  # breaks = bk,
  main='BMC correction')
```
#limma
```{r}
#limma
ps_limma<- ps_all
otu_table(ps_limma) <- otu_table(panc.limma,taxa_are_rows = F)
ps_obj <- ps_limma
ps_obj_meta <- sample_data(ps_obj)
cols_of_interest =c('group','batch')
meta_for_heatmap <- ps_obj_meta[, (names(ps_obj_meta) %in% cols_of_interest)]
meta_for_heatmap <- as(meta_for_heatmap, 'data.frame')
ordered_meta <- meta_for_heatmap[order(meta_for_heatmap$group,meta_for_heatmap$batch),]
pheatmap( 
  t(ps_limma@otu_table), 
  cluster_rows = F, 
  cluster_cols = T,
  fontsize_row = 4, fontsize_col = 5,
  fontsize = 12,
  clustering_distance_rows = 'euclidean',
  clustering_method = 'ward.D',
  treeheight_row = 15,
  annotation_col = meta_for_heatmap,
  annotation_colors = annotation_colors,
  height = 40,
  # col= hmcols, 
  # breaks = bk,
  main='Limma correction')
```
#svd
```{r}
#svd
ps_svd<- ps_all
otu_table(ps_svd) <- otu_table(panc.svd,taxa_are_rows = F)
ps_obj <- ps_svd
ps_obj_meta <- sample_data(ps_obj)
cols_of_interest =c('group','batch')
meta_for_heatmap <- ps_obj_meta[, (names(ps_obj_meta) %in% cols_of_interest)]
meta_for_heatmap <- as(meta_for_heatmap, 'data.frame')
ordered_meta <- meta_for_heatmap[order(meta_for_heatmap$group,meta_for_heatmap$batch),]
pheatmap( 
  t(ps_svd@otu_table), 
  cluster_rows = F, 
  cluster_cols = T,
  fontsize_row = 4, fontsize_col = 5,
  fontsize = 12,
  #clustering_distance_rows = 'euclidean',
  #clustering_method = 'ward.D',
  treeheight_row = 15,
  annotation_col = meta_for_heatmap,
  annotation_colors = annotation_colors,
  height = 40,
  # col= hmcols, 
  # breaks = bk,
  main='svd correction')
```
#percentile
```{r}
#percentile
ps_pn<- ps_all
otu_table(ps_pn) <- otu_table(panc.percentile ,taxa_are_rows = F)
ps_obj <- ps_pn
#ps_obj <- filter_taxa(ps_genus, function(x){sum(x > 0) > 10}, prune = TRUE)
# prepare metadata
ps_obj_meta <- sample_data(ps_obj)
cols_of_interest =c('group','batch')
meta_for_heatmap <- ps_obj_meta[, (names(ps_obj_meta) %in% cols_of_interest)]
meta_for_heatmap <- as(meta_for_heatmap, 'data.frame')
ordered_meta <- meta_for_heatmap[order(meta_for_heatmap$group,meta_for_heatmap$batch),]
#limma.scale <- scale(ps_limma@otu_table, center = T, scale = T) 
# scale on samples
#limma.scale <- scale(t(limma.scale), center = T, scale = T) 
#dev.new(width=20, height=20)
pheatmap( 
  t(ps_pn@otu_table), 
  cluster_rows = F, 
  cluster_cols = T,
  fontsize_row = 4, fontsize_col = 5,
  fontsize = 12,
  #clustering_distance_rows = 'euclidean',
  #clustering_method = 'ward.D',
  treeheight_row = 15,
  annotation_col = meta_for_heatmap,
  annotation_colors = annotation_colors,
  height = 40,
  # col= hmcols, 
  # breaks = bk,
  main='PN correction')


```
#Fabatch
```{r}
ps_fa<- ps_all
otu_table(ps_fa) <- otu_table(panc.fabatch ,taxa_are_rows = F)
ps_obj <- ps_fa
#ps_obj <- filter_taxa(ps_genus, function(x){sum(x > 0) > 10}, prune = TRUE)
# prepare metadata
ps_obj_meta <- sample_data(ps_obj)
cols_of_interest =c('group','batch')
meta_for_heatmap <- ps_obj_meta[, (names(ps_obj_meta) %in% cols_of_interest)]
meta_for_heatmap <- as(meta_for_heatmap, 'data.frame')
ordered_meta <- meta_for_heatmap[order(meta_for_heatmap$group,meta_for_heatmap$batch),]
pheatmap( 
  t(panc.fabatch), 
  cluster_rows = F, 
  cluster_cols = T,
  fontsize_row = 4, fontsize_col = 5,
  fontsize = 12,
  #clustering_distance_rows = 'euclidean',
  #clustering_method = 'ward.D',
  treeheight_row = 15,
  annotation_col = meta_for_heatmap,
  annotation_colors = annotation_colors,
  height = 40,
  # col= hmcols, 
  # breaks = bk,
  main='fabatch correction')

```

```{r}
panc.b.coeff <- c()
for(i in 1:ncol(panc.clr)){
  res <- lm(panc.clr[ ,i] ~ panc.gr + panc.batch)
  sum.res <- summary(res)
  panc.b.coeff[i] <- sum.res$coefficients[3,1]
}
par(mfrow = c(2,2))
hist(panc.b.coeff,col = 'gray')
plot(density(panc.b.coeff))
qqnorm(panc.b.coeff)
qqline(panc.b.coeff, col='red')
par(mfrow = c(1,1))



```
#The pRDA method is a multivariate method to assess globally the effect of batch and treatment (two separate models)
```{r}

panc.data.design <- numeric()
panc.data.design$group <- panc.gr
panc.data.design$batch <- panc.batch

# before
# conditioning on a batch effect
panc.rda.before1 <- rda(panc.clr ~ group + Condition(batch), 
                          data = panc.data.design)
panc.rda.before2 <- rda(panc.clr ~ batch + Condition(group), 
                          data = panc.data.design)

# amount of variance
panc.rda.bat_prop.before <- panc.rda.before1$pCCA$tot.chi*100/panc.rda.before1$tot.chi
panc.rda.trt_prop.before <- panc.rda.before2$pCCA$tot.chi*100/panc.rda.before2$tot.chi

# BMC
# conditioning on a batch effect
panc.rda.bmc1 <- rda(panc.bmc ~ group + Condition(batch), 
                       data = panc.data.design)
panc.rda.bmc2 <- rda(panc.bmc ~ batch + Condition(group), 
                       data = panc.data.design)

# amount of variance
panc.rda.bat_prop.bmc <- panc.rda.bmc1$pCCA$tot.chi*100/panc.rda.bmc1$tot.chi
panc.rda.trt_prop.bmc <- panc.rda.bmc2$pCCA$tot.chi*100/panc.rda.bmc2$tot.chi


# combat
# conditioning on a batch effect
panc.rda.combat1 <- rda(panc.combat ~ group + Condition(batch), 
                          data = panc.data.design)
panc.rda.combat2 <- rda(panc.combat ~ batch + Condition(group), 
                          data = panc.data.design)

# amount of variance
panc.rda.bat_prop.combat <- panc.rda.combat1$pCCA$tot.chi*100/panc.rda.combat1$tot.chi
panc.rda.trt_prop.combat <- panc.rda.combat2$pCCA$tot.chi*100/panc.rda.combat2$tot.chi


# limma
# conditioning on a batch effect
panc.rda.limma1 <- rda(panc.limma ~ group + Condition(batch), 
                         data = panc.data.design)
panc.rda.limma2 <- rda(panc.limma ~ batch + Condition(group), 
                         data = panc.data.design)

# amount of variance
panc.rda.bat_prop.limma <- panc.rda.limma1$pCCA$tot.chi*100/panc.rda.limma1$tot.chi
panc.rda.trt_prop.limma <- panc.rda.limma2$pCCA$tot.chi*100/panc.rda.limma2$tot.chi


# percentile
# conditioning on a batch effect
panc.rda.percentile1 <- rda(panc.percentile ~ group + Condition(batch), 
                              data = panc.data.design)
panc.rda.percentile2 <- rda(panc.percentile ~ batch + Condition(group), 
                              data = panc.data.design)

# amount of variance
panc.rda.bat_prop.percentile <- panc.rda.percentile1$pCCA$tot.chi*100/panc.rda.percentile1$tot.chi
panc.rda.trt_prop.percentile <- panc.rda.percentile2$pCCA$tot.chi*100/panc.rda.percentile2$tot.chi


# SVD
# conditioning on a batch effect
panc.rda.svd1 <- rda(panc.svd ~ group + Condition(batch), 
                       data = panc.data.design)
panc.rda.svd2 <- rda(panc.svd ~ batch + Condition(group), 
                       data = panc.data.design)

# amount of variance
panc.rda.bat_prop.svd <- panc.rda.svd1$pCCA$tot.chi*100/panc.rda.svd1$tot.chi
panc.rda.trt_prop.svd <- panc.rda.svd2$pCCA$tot.chi*100/panc.rda.svd2$tot.chi


# Fabatch
# conditioning on a batch effect
panc.rda.fabatch1 <- rda(panc.fabatch ~ group + Condition(batch), 
                     data = panc.data.design)
panc.rda.fabatch2 <- rda(panc.fabatch ~ batch + Condition(group), 
                     data = panc.data.design)

# amount of variance
panc.rda.bat_prop.fabatch <- panc.rda.fabatch1$pCCA$tot.chi*100/panc.rda.fabatch1$tot.chi
panc.rda.trt_prop.fabatch <- panc.rda.fabatch2$pCCA$tot.chi*100/panc.rda.fabatch2$tot.chi



#We now represent the amount of variance explained by batch and treatment estimated with pRDA:
  
  # proportion
panc.rda.prop.before <- c(panc.rda.bat_prop.before, 
                          panc.rda.trt_prop.before)
panc.rda.prop.bmc <- c(panc.rda.bat_prop.bmc, 
                       panc.rda.trt_prop.bmc)
panc.rda.prop.combat <- c(panc.rda.bat_prop.combat, 
                          panc.rda.trt_prop.combat)
panc.rda.prop.limma <- c(panc.rda.bat_prop.limma, 
                         panc.rda.trt_prop.limma)
panc.rda.prop.percentile <- c(panc.rda.bat_prop.percentile, 
                              panc.rda.trt_prop.percentile)
panc.rda.prop.svd <- c(panc.rda.bat_prop.svd, 
                       panc.rda.trt_prop.svd)
panc.rda.prop.fabatch <- c(panc.rda.bat_prop.fabatch, 
                       panc.rda.trt_prop.fabatch)

# merge results
panc.rda.prop.val <- c(panc.rda.prop.before, panc.rda.prop.bmc, 
                       panc.rda.prop.combat, panc.rda.prop.limma, 
                       panc.rda.prop.percentile, panc.rda.prop.svd, panc.rda.prop.fabatch)

# add batch, trt and method info
panc.rda.prop <- data.frame(prop = panc.rda.prop.val, 
                              prop.r = round(panc.rda.prop.val, 2), 
                              Method = rep(c('Before', 'BMC', 'ComBat', 
                                             'rBE', 'PN', 'SVD','Fabatch'), each = 2), 
                              Type = rep(c('Batch', 'Group'), 7))

# reorder levels
panc.rda.prop$Method <- factor(panc.rda.prop$Method, 
                                 levels = unique(panc.rda.prop$Method))

ggplot(data = panc.rda.prop, aes(x = Method, y = prop, fill = Type)) + 
  geom_bar(stat = "identity", position = 'dodge', colour = 'black') + 
  geom_text(data = panc.rda.prop, aes(Method, prop + 2.5, label = prop.r), 
            position = position_dodge(width = 0.9), size = 3) + theme_bw() + 
  labs(y = "Variance explained (%)") + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1), 
        panel.grid = element_blank(), axis.text = element_text(size = 12), 
        axis.title = element_text(size = 15), legend.title = element_text(size = 15),
        legend.text = element_text(size = 12)) + ylim(0,100)

```


